---
title: "Programming in R"
subtitle: Task 3
author: ' Sofía Gianelli Nan, Gabriela Levenfeld Sabau and Miguel Díaz-Plaza Cabrera'
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: TRUE
    toc_depth: 6
    keep_tex: TRUE
    includes:
        in_header: "wrap-code.tex"
  geometry: "left=3cm,right=3cm,top=2cm,bottom=2cm"
  mathjax: local
  self_contained: false
font-family: Helvetica

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\pagebreak

# 1. Exercise I: Create the following vectors

-   **a)  [10, 11, 12, . . . , 38]**

```{r ex1_a}
vect_a <- seq(10, 38, 1)
vect_a
```


-   **b) [30, 29, 28, . . . , 1]**

```{r ex1_b}
vect_b <- c(30:1)
vect_b
```


-   **c) [1, 2, 3, 4, 3, 2, 1]**

```{r ex1_c}
vect_c <- c(seq(1, 4, 1), seq(3, 1, -1))
vect_c
```


-   **d) [2, 4, 6, . . . , 16, 18, 20]**

```{r ex1_d}
vect_d <- seq(2, 20, 2)
vect_d
```


-   **e) [1, 2, 3, 1, 2, 3, . . . , 1, 2, 3]**

```{r ex1_e}
vect_e <- rep(seq(1, 3, 1), 10)
vect_e
```


-   **f) [1, 2, 3, 1, 2, 3, . . . , 1]**

```{r ex1_f}
vect_f <- head(vect_e, -2)
vect_f
```


-   **g) ["label 1", "label 2",. . . , "label 30"]**

```{r ex1_g}
vect_g <- paste('label', seq(1, 30))
vect_g
```


-   **h) ["label-1", "label-2",. . . , "label-30"]**

```{r ex1_h}
vect_h <- paste('label', seq(1, 30), sep = '-')
vect_h
```


-   **i) $x^2e^x$, x = 0.1, 0.2, ..., 1.**

```{r ex1_i}
x <- seq(0.1, 1, 0.1)
vect_i <- (x^2)*(exp(x))
vect_i
```

\pagebreak

# 2. Exercise II: Calculate the followings sums with and without using loops

-   **$\sum_{{j=5}}^{{23}} (j^2 + 3*j^{0.5})$**

```{r ex2_a}
# With loops
j <- 5; sol1 = 0;
while(j<=23){
  sol1 = sol1 + (j^2 + 3*(j^0.5))
  j = j+1
}
sol1

# Without loops
j_vect <- 5:23; result1 <- 0;
f_j_values <- function(x, y){
  (x^2 + 3*(x^0.5))
}
result1 <- sum(outer(j_vect, 0, FUN = f_j_values))
result1
```



-   **$\sum_{{i=1}}^{{18}} \frac{1.3^i}{i}$**

```{r ex2_b}
# With loops
i <- 1; sol2 <-0;
while(i <= 18){
  sol2 = sol2 + (1.3^i)/i
  i = i+1
}
sol2

# Without loops
i_vect <- 1:18; result2 <- 0;
f_i_values <- function(x, y){
  ((1.3^x)/x)
}
result2 <- sum(outer(i_vect, 0, FUN = f_i_values))
result2
```



-   **$\sum_{i=1}^{10}\sum_{j=1}^{6} \frac{i^4}{3+j}$**

```{r ex2_c}
# With loops
sol3 <- 0;
for(i in 1:10){
  for(j in 1:6){
    sol3 = sol3 + (i^4)/(3+j)
  }
}
sol3

# Without loops
i_vect <- 1:10; j_vect <- 1:6; result3 <- 0;
f_ij_values <- function(x, y){
  (x^4)/(3+y)
}
result3 <- sum(outer(i_vect, j_vect, FUN = f_ij_values))
result3
```

\pagebreak

# 3. Exercise III

**(a) What does the next code do?**

```{r ex3_a}
set.seed(75)
M = matrix(sample(1:10, size=60, replace=TRUE), nrow=6, ncol=10)
```


The first line of code allows us to create the same randoms numbers every time we execute the code. Then we create a matrix named M which has 6 rows and 10 columns thanks to the last two parameters. Finally, the 60 numbers inside the matrix are selected randomly from 1 to 10 and by setting "replace=TRUE", it is being specify that the same number can be chosen several times.



**(b) Find the number of entries in each row that are greater than 4.**

```{r ex3_b}
greater_4 <- function(row){
  sum(row>4)
}
apply(M, MARGIN = 1, FUN = greater_4)
```


Where first position of the output corresponds to the number of values greater than 4 in the first row, the second one to the second row and so on.



**(c) Replace the third column of the previous vector M by the sum of the second and third column.**

```{r ex3_c}
M[,3] = M[,2]+M[,3]
M
```

\pagebreak

# 4. Exercise IV

**Write a function which takes a single argument which is a matrix. The function must return a matrix which is the same as the function argument but every odd number is doubled.**

```{r ex4}
oddDoubled <- function(X){
  ifelse(X %% 2 == 0, X, 2*X)
}
a <- matrix(1:8, nrow = 2, ncol = 4)
newA = oddDoubled(a)
```


```{r}
a
```


```{r}
newA
```


To illustrated that this code works, it is attached an example.
